\LaTeXe\ guide for the Journal of Functional Programming

### 2019-05-03 papers/jfp/7-experience.tex
Rework polymorphic values
Tweaks to S7

There are two main
aspects to the original \Shake paper~\cite{mitchell2012shake} that are described
more clearly in this paper. Firstly, the rebuilder can now be described using
verifying step traces~\S\ref{sec-step-traces}, with a much clearer relationship
to the unoptimised verifying traces of~\S\ref{sec-verifying-traces}. Secondly,
in the original paper the tasks (there called ``actions'')
were described in continuation-passing style using this data type\footnote{
The original paper uses concrete types \hs{Key} and \hs{Value}. Here we
generalise these types to \hs{k} and \hs{v}, and also add \hs{a} so that
\hs{Action}~\hs{k}~\hs{v} can be an instance of \hs{Monad}}.
\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
data Action k v a = Finished a
                  | Depends k (v -> Action k v a)
\end{minted}
\vspace{1mm}
\noindent
In this paper we describe tasks more directly, in a monadic (or applicative or functorial)
style. But in fact the two are equivalent: \hs{Task Monad k v} is
isomporphic to \hs{Action k v v}.
To be concrete, the functions \hs{toAction} and \hs{fromAction} witness the
isomorphism in both directions.
\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
instance Monad (Action k v) where
    return = Finished
    Finished x    >>= f = f x
    Depends ds op >>= f = Depends ds $ \v -> op v >>= f

@@

\noindent
Similarly, in the original paper \Make tasks were described as equivalent to

@@

\noindent
These functions are suspiciously like \hs{recordCT} and \hs{constructCT}
from~\S\ref{sec-constructive-traces}, with their differences perhaps the most
illustrative of the changes required\footnote{
  We have made some minor changes from
actual \Shake, like replacing \hs{Key} for \hs{k}, to reduce irrelevant
differences.}.

@@

Interestingly, although \Shake is not a self-tracking build
system~(\S\ref{sec-tracking-aspects}), \Hadrian implements a little
domain-specific language for constructing build command lines, and then tracks
command lines by treating them as a type of values -- an example of
\emph{partial self-tracking} made possible by \Shake's support for value
polymorphism~(\S\ref{sec-polymorphism}).

The subsequent development of \Cloud \Shake allows GHC developers to take advantage of
sharing build results between builds, which is particularly promising in the
context of GHC's continuous integration (CI) infrastructure responsible for
building and testing every GHC commit. Building GHC~8.8 from scratch with
default settings takes around 1 hour on Windows using both \Hadrian and the
original \Make-based build system. This includes building the compiler itself,
29~bundled libraries, such as \cmd{base} (each in vanilla and profiled way), and
6~bundled executables, such as \cmd{Haddock}. A single commit typically modifies
only a small part of the whole codebase, making the cloud-build functionality
highly desirable. For example, if a commit does not touch any GHC source files
(e.g. only modifies the documentation), \Hadrian can build GHC from scratch in under
3~minutes, by simply creating symbolic links to the previously build results
stored in the \Cloud~\Shake cache.

A small number of GHC build rules cannot be cached. These rules register new libraries
in the GHC package database, and rather than producing a output file, they mutate
a shared file. \Cloud~\Shake provides a way to manually label such
build rules, thereby excluding them from caching.

>>>>

There are two main
aspects to the original \Shake paper~\cite{mitchell2012shake} that are described
more clearly in this paper. Firstly, the rebuilder can now be described using
verifying step traces~\S\ref{sec-step-traces}, with a much clearer relationship
to the unoptimised verifying traces of~\S\ref{sec-verifying-traces}. Secondly,
in the original paper the tasks (there called ``actions'')
were described in continuation-passing style using the data type:\footnote{
The original paper uses concrete types \hs{Key} and \hs{Value}. Here we
generalise these types to \hs{k} and \hs{v}, and also add \hs{a} so that
\hs{Action}~\hs{k}~\hs{v} can be an instance of \hs{Monad}}
\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
data Action k v a = Finished a
                  | Depends k (v -> Action k v a)
\end{minted}
\vspace{1mm}
\noindent
In this paper we describe tasks more directly, in a monadic (or applicative or functorial)
style. But in fact the two are equivalent: \hs{Task Monad k v} is
isomporphic to \hs{Action k v v}.
To be concrete, the functions \hs{toAction} and \hs{fromAction} witness the
isomorphism in both directions.
\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
instance Monad (Action k v) where
    return = Finished
    Finished x    >>= f = f x
    Depends ds op >>= f = Depends ds $ \v -> op v >>= f

@@

\noindent
Similarly, in the original paper \Make tasks were described as:

@@

\noindent
These functions are suspiciously like \hs{recordCT} and \hs{constructCT}
from~\S\ref{sec-constructive-traces}, with their differences perhaps the most
illustrative of the changes required\footnote{We have made some minor changes from
actual \Shake, like replacing \hs{Key} for \hs{k}, to reduce irrelevant
differences.}.

@@

Interestingly, although \Shake is not a self-tracking build
system~(\S\ref{sec-tracking-aspects}), \Hadrian implements a little
domain-specific language for constructing build command lines, and then tracks
command lines by treating them as a type of values -- an example of
\emph{partial self-tracking} made possible by \Shake's support for key-dependent
value types~(\S\ref{sec-polymorphism}).

The subsequent development of \Cloud \Shake allows GHC developers to take advantage of
sharing build results between builds, which is particularly promising in the
context of GHC's continuous integration (CI) infrastructure responsible for
building and testing every GHC commit. Building GHC~8.8 from scratch with
default settings takes around 1 hour on Windows using either \Hadrian or the
original \Make-based build system. This time includes building the compiler itself,
29~bundled libraries, such as \cmd{base} (each in vanilla and profiled way), and
6~bundled executables, such as \cmd{Haddock}. A single commit typically modifies
only a small part of the whole codebase, making the cloud-build functionality
highly desirable. For example, if a commit does not touch any GHC source files
(e.g. only modifies the documentation), \Hadrian can build GHC from scratch in under
3~minutes, by simply creating symbolic links to the previously build results
stored in the \Cloud~\Shake cache.

A small number of GHC build rules cannot be cached. These rules register new libraries
in the GHC package database, and rather than producing a output file, they mutate
a shared file. \Cloud~\Shake provides a way to manually label such
build rules to exclude them from caching.
